---
title: Week 03 Lab Learning Notes -- 'PostgreSQL' Queries for Joining Tables
author: Owen
date: 2024-05-21
date-format: "dddd MMM D, YYYY"

format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    number-sections: true
    number-depth: 5
---

```{r}
#| label: setup
#| include: false

library(knitr)
library(kableExtra)
library(DBI)
con <- dbConnect(odbc::odbc(), dsn="investments", timeout = 10)


```
## Inner Joins
### Inner Joins for One-to-One Relationships
Here we have 2 tables

```{r echo=FALSE}
query <- "SELECT * FROM shareholders;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "Table 1: Shareholder information")

query <- "SELECT * FROM traded_companies;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "Table 2: Companies traded on the stock market")

query <- "SELECT
    sh.sh_holder_id,
    sh.sh_holder_name,
    tc.comp_id
FROM shareholders AS sh
INNER JOIN traded_companies AS tc
    ON sh.sh_holder_name = tc.comp_name
ORDER BY sh.sh_holder_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "2 records")
```



### Inner Joins for One-to-Many Relationships
We have another table

```{r echo=FALSE}
query <- "SELECT * FROM products;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "Table 3: Information on products sold by the companies in the traded_companies table")
```
```{r echo=FALSE}
query <- "SELECT
    tc.comp_name,
    tc.comp_id,
    pr.prod_name
FROM traded_companies AS tc
INNER JOIN products AS pr
    ON tc.comp_id = pr.comp_id
ORDER BY tc.comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "7 records")
```
### Inner Joins for Many-to-Many Relationships
```{r echo=FALSE}
query <- "SELECT
    sh.sh_holder_name,
    sh.country,
    tc.comp_name
FROM shareholders AS sh
INNER JOIN traded_companies AS tc
    ON sh.country = tc.country
ORDER BY sh.sh_holder_id, tc.comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "9 records")
```
## Outer Joins
### Left Joins
```{r echo=FALSE}
query <- "SELECT
    sh.sh_holder_name,
    sh.country,
    tc.comp_name
FROM shareholders AS sh
LEFT JOIN traded_companies AS tc
    ON sh.country = tc.country
ORDER BY sh.sh_holder_id, tc.comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "11 records")
```
### Right Joins
```{r echo=FALSE}
query <- "SELECT
    sh.sh_holder_name,
    tc.country,
    tc.comp_name
FROM shareholders AS sh
RIGHT JOIN traded_companies AS tc
    ON sh.country = tc.country
ORDER BY sh.sh_holder_id, tc.comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "11 records")
```
### Full Outer Joins
```{r echo=FALSE}
query <- "SELECT
    sh.sh_holder_name,
    tc.comp_name,
    COALESCE(sh.country, tc.country) AS country
FROM shareholders AS sh
FULL OUTER JOIN traded_companies AS tc
    ON sh.country = tc.country
ORDER BY sh.sh_holder_id, tc.comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "14 records")
```
## Filtering Joins
### Filter Joins
```{r echo=FALSE}
query <- "SELECT
    sh_holder_name,
    country
FROM shareholders
WHERE country IN (
    SELECT country
    FROM traded_companies
)
ORDER BY sh_holder_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "6 records")
```
### Anti-Joins
```{r echo=FALSE}
query <- "SELECT
    sh_holder_name,
    country
FROM shareholders
WHERE country NOT IN (
    SELECT country
    FROM traded_companies
)
ORDER BY sh_holder_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "2 records")
```
## Exercises — Switching Tables in Filtering Joins
a. In [Section 3.1](### Filter Joins), a semi-join was performed based on the country column with shareholders as the left table and traded_companies as the right table. What would happen if the tables were switched?

Answer: it will produced a different table

Result:
```{r echo=FALSE}
query <- "SELECT
    comp_id,
    country
FROM traded_companies
WHERE country IN (
    SELECT country
    FROM shareholders
)
ORDER BY comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "6 records")
```
b. Repeat the exercise for the anti-join in [Section 3.2](### Anti-Joins)

Result:
```{r echo=FALSE}
query <- "SELECT
    comp_id,
    country
FROM traded_companies
WHERE country NOT IN (
    SELECT country
    FROM shareholders
)
ORDER BY comp_id;"

data <- dbGetQuery(con, query)
knitr::kable(data,caption = "3 records")
```



