---
title: Week 04 Lab Lab Notes -- Sample 'PostgreSQL' questions Queries for Joining Tables
author: Owen
date: 2024-05-24
date-format: "dddd MMM D, YYYY"

format: 
  html:
    toc: true
    toc-depth: 5
    toc-location: left
    number-sections: true
    number-depth: 5
---

```{r}
#| label: setup
#| include: false

library(knitr)
library(kableExtra)
library(DBI)
con <- dbConnect(odbc::odbc(), dsn="NYC_db", timeout = 10)


```
```{r}
#| label: setup
#| include: false

# Install necessary packages
install.packages("dplyr")
install.packages("ggplot2")

# Load necessary libraries
library(ggplot2)
library(dplyr)

# Create the dataframe
airports_df <- data.frame(
  faa = c("04G", "06A", "06C", "06N"),
  name = c("Lansdowne Airport", "Moton Field Municipal Airport", "Schaumburg Regional", "Randall Airport"),
  lat = c(41.1304722, 32.4605722, 41.9893408, 41.431912),
  lon = c(-80.6195833, -85.6800278, -88.1012428, -74.3915611),
  alt = c(1044, 264, 801, 523),
  tz = c(-5, -6, -6, -5),
  dst = c("A", "A", "A", "A"),
  tzone = c("America/New_York", "America/Chicago", "America/Chicago", "America/New_York")
)

# Calculate the frequency of each airport appearing
airport_freq <- airports_df %>%
  group_by(name) %>%
  summarise(count = n())

# Add a column with labels for the pie chart
airport_freq <- airport_freq %>%
  mutate(label = paste(name, "(", count, ")", sep = ""))

# Plot the pie chart
ggplot(airport_freq, aes(x = "", y = count, fill = name)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  labs(title = "Frequency of Airports Appearing") +
  theme_void()

```
1. List all flights that were delayed more than 30 minutes on departure.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT *
FROM flights
WHERE dep_delay > 30;
```

2. Retrieve the flight number, carrier name, and destination airport for all flights that departed from EWR (Newark Liberty International Airport).

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT f.flight, a.name, f.dest
FROM flights f
JOIN airlines a ON f.carrier = a.carrier
WHERE origin = 'EWR';

```

3. Find the average arrival delay for each carrier.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT carrier, AVG(arr_delay) AS avg_arr_delay
FROM flights
GROUP BY carrier;
```

4. List the tail number, manufacturer, and model of planes that have flights departing from LGA (LaGuardia Airport).

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT DISTINCT p.tailnum, p.manufacturer, p.model
FROM flights f
JOIN planes p ON f.tailnum = p.tailnum
WHERE f.origin = 'LGA';
```

5. Retrieve the flight number, origin, destination, and the name of the destination airport for flights that traveled more than 1000 miles.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT f.flight, f.origin, f.dest, a.name AS destination_airport_name
FROM flights f
JOIN airports a ON f.dest = a.faa
WHERE f.distance > 1000;
```

6. Find all planes that are in the planes table but did not have any recorded flights in the flights table.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT p.tailnum, p.manufacturer, p.model
FROM planes p
LEFT JOIN flights f ON p.tailnum = f.tailnum
WHERE f.tailnum IS NULL;
```

7. For each day, count the number of flights that were scheduled to depart.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT DATE_TRUNC('day', time_hour) AS day, COUNT(*) AS num_flights
FROM flights
GROUP BY DATE_TRUNC('day', time_hour)
ORDER BY day;
```

8. List the flights along with their corresponding weather conditions at the origin airport for January 1, 2013.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT f.flight, f.origin, w.temp, w.precip, w.wind_speed
FROM flights f
JOIN weather w ON f.origin = w.origin AND f.year = w.year AND f.month = w.month AND f.day = w.day AND f.hour = w.hour
WHERE f.year = 2013 AND f.month = 1 AND f.day = 1;
```

9.  Find the carriers that have an average departure delay of more than 10 minutes. Find the carriers that have an average departure delay of more than 10 minutes.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT carrier, AVG(dep_delay) AS avg_dep_delay
FROM flights
GROUP BY carrier
HAVING AVG(dep_delay) > 10;
```

10. Retrieve the flight number, departure time, scheduled departure time, and manufacturer of planes that departed with a delay of more than 60 minutes.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT f.flight, f.dep_time, f.sched_dep_time, p.manufacturer
FROM flights f
JOIN planes p ON f.tailnum = p.tailnum
WHERE f.dep_delay > 60;
```

11. List all flights that departed after the average departure time on January 1, 2013.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT *
FROM flights
WHERE dep_time > (
    SELECT AVG(dep_time)
    FROM flights
    WHERE year = 2013 AND month = 1 AND day = 1
) AND year = 2013 AND month = 1 AND day = 1;
```

12. Retrieve the names of carriers that had flights departing from EWR with a delay greater than the average departure delay of all flights from EWR.

```{sql}
#| connection: con
#| eval: true
#| echo: true

WITH avg_dep_delay_ewr AS (
    SELECT AVG(dep_delay) AS avg_dep_delay
    FROM flights
    WHERE origin = 'EWR'
)
SELECT DISTINCT a.name AS carrier_name
FROM flights f
JOIN avg_dep_delay_ewr avg ON f.dep_delay > avg.avg_dep_delay
JOIN airlines a ON f.carrier = a.carrier
WHERE f.origin = 'EWR';

```

13. List the flight number, origin, and the maximum departure delay for flights on the same day.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT flight, origin, MAX(dep_delay) AS max_dep_delay
FROM flights
GROUP BY year, month, day, flight, origin;
```


14. Find all flights that had a departure delay greater than the average departure delay of their respective carrier.

```{sql}
#| connection: con
#| eval: true
#| echo: true

WITH avg_dep_delay_per_carrier AS (
    SELECT carrier, AVG(dep_delay) AS avg_dep_delay
    FROM flights
    GROUP BY carrier
)
SELECT f.*
FROM flights f
JOIN avg_dep_delay_per_carrier ad
ON f.carrier = ad.carrier
WHERE f.dep_delay > ad.avg_dep_delay;

```
15. For each carrier, find the average arrival delay and list carriers that have an average arrival delay of more than 10 minutes.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT carrier, avg_arr_delay
FROM (
    SELECT carrier, AVG(arr_delay) AS avg_arr_delay
    FROM flights
    GROUP BY carrier
) AS avg_delays
WHERE avg_arr_delay > 10;
```

16. List the tail numbers of planes that have flown more than the average distance of all flights.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT DISTINCT tailnum
FROM flights
WHERE distance > (
    SELECT AVG(distance)
    FROM flights
);
```

17. Find all flights that have corresponding weather data with a temperature higher than 80°F at the origin airport.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT *
FROM flights f
WHERE EXISTS (
    SELECT 1
    FROM weather w
    WHERE f.origin = w.origin AND f.year = w.year AND f.month = w.month AND f.day = w.day AND w.temp > 80
);
```

18. List carriers that have more than 100 flights with departure delays.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT carrier, COUNT(*) AS num_flights
FROM flights
GROUP BY carrier
HAVING COUNT(*) > 100 AND COUNT(dep_delay) > (
    SELECT COUNT(*)
    FROM flights
    WHERE dep_delay IS NOT NULL
);
```

19. Find all planes that have not flown to JFK (John F. Kennedy International Airport).

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT tailnum
FROM planes
WHERE tailnum NOT IN (
    SELECT DISTINCT tailnum
    FROM flights
    WHERE dest = 'JFK'
);
```

20. List each flight along with the average departure delay for flights from the same origin on the same day.

```{sql}
#| connection: con
#| eval: true
#| echo: true

SELECT flight,
       origin,
       year,
       month,
       day,
       AVG(dep_delay) AS avg_dep_delay
FROM flights
GROUP BY origin, year, month, day, flight;

```


